Efficient implementation of penalized models and their application to genetic risk prediction

### 2018-05-08 paper/draft.tex
some modifs by H

\title{Highly efficient implementation of penalized models and their application to genetic risk prediction}
\author{Florian Priv\'e\,$^{\text{ 1,}*}$, Hugues Aschard\,$^{\text{2}}$ and Michael G.B. Blum\,$^{\text{1,}*}$}

@@

\abstract{
Polygenic Risk Scores (PRSs) consist in combining the information across many single-nucleotide polymorphisms (SNPs) in a score reflecting the genetic risk of developing a disease. 
PRSs might have a major public health impact, possibly allowing for screening campaigns to identify high-genetic risk individuals for a given disease.  The most commonly used method to derive PRSs is called ``Clumping+Thresholding'' (C+T). It uses only univariate genome-wide association studies (GWAS) summary statistics, which makes it very suitable and fast. However, these PRS rely on non-optimal heuristics.

In R package bigstatsr, we implemented linear and logistic regression models that can be used to jointly estimate SNP effects for computing PRS. We use an efficient algorithm with a new implementation in order for this model to be used on huge datasets.

In this paper, we present a comprehensive comparative study of the C+T method, our penalized logistic regression and a derivation of random forests. The penalized logistic regression consistently achieves higher predictive performance than the two other methods while being faster than the C+T method. 
Moreover, predictive performance of the C+T method is very sensitive to the threshold of inclusion of SNPs, whose optimal value depends on disease architecture. 
The improvement in predictive performance is more pronounced when there are few effects located in nearby genomic regions with correlated SNPs, and with an increasing sample size. 
When analyzing real SNP data from a case-control study for celiac disease, penalized logistic regression provides an AUC of 89\% while the standard C+T method provides a smaller AUC of 82\%.

[TOMOD] Instead of computing polygenic risk scores based on GWAS univariate summary statistics, we recommend to use penalized logistic regression as implemented in the R package bigstatsr to obtain more accurate polygenic risk scores when individual genotype data are available.\\

@@

Many methods have been developed to predict disease status, or more generally any phenotype, based on SNP information.
A commonly used method, called ``P+T'' or ``C+T'' (which stands for ``Clumping and Thresholding'') -- or even just PRS -- is used to derive PRSs from results of Genome-Wide Association Studies (GWASs) \cite[]{chatterjee2013projecting,dudbridge2013power,evans2009harnessing,purcell2009common,wray2007prediction}. 
This technique uses GWAS summary statistics only, which makes it very convenient to compute and also very fast.
Linear Mixed-Models (LMMs) are another widely-used method in fields such as plant and animal breeding or for predicting highly heritable quantitative human phenotypes such as height \cite[]{lello2017accurate,yang2010common}. 
Yet, models resulting from LMM, known e.g.\ as ``gBLUP'', are not optimal for predicting disease status based on genotypes \cite[]{abraham2013performance}. 
Moreover, these methods and their derivatives are often computationally demanding, both in terms of memory and time requirements, which makes them unlikely to be used for prediction on very large datasets \cite[]{golan2014effective,maier2015joint,speed2014multiblup,zhou2013polygenic}.
Finally, statistical learning methods have also been used to derive PRSs for complex human diseases by jointly estimating SNP effects. Such methods include joint logistic regression, Support Vector Machine (SVM) and random forests \cite[]{abraham2012sparsnp,abraham2014accurate,botta2014exploiting,okser2014regularized,wei2009disease}.

We recently developed two R packages, bigstatsr and bigsnpr, for efficiently analyzing large-scale genome-wide data \cite[]{prive2017efficient}. Package bigstatsr includes an efficient algorithm with a new implementation for computing sparse linear and logistic regressions on huge datasets as large as the UK Biobank \cite[]{bycroft2017genome}.
In this paper, we present a comprehensive comparative study of the C+T method, our implementation of penalized logistic regression and the T-Trees algorithm, a derivation of random forests that has shown high predictive performance \cite[]{botta2014exploiting}. 
Note that in this comparison we do not include any LMM method for the reasons mentioned before and do not include any SVM method because it is expected to give similar results to logistic regression \cite[]{abraham2012sparsnp}.
For the C+T model, we compare different thresholds of inclusion of SNPs.
For the penalized logistic regression, we include two practical improvements. 
First, we introduce a procedure that we call Cross-Model Selection and Averaging (CMSA) for automatically choosing the regularization parameter of the penalized logistic regression, which directly affects the number of SNPs included in the model. 
Secondly, we also show how to use feature engineering in order to capture not only linear effects, but also recessive and dominant effects.

There have already been comparative studies on predicting human binary phenotypes based on genotype data. 
\cite{abraham2013performance} showed that sparse penalized methods performed better than the standard C+T method and LMMs, for a wide-range of human diseases.
\cite{zhou2013polygenic} used simulations on real genotypes of Australian individuals to show that a hybrid of linear mixed models and sparse regression models performed better in a range of sparse to polygenic disease architectures. 
\cite{spiliopoulou2015genomic} investigated the effect of relatedness in prediction performance.
\cite{ware2017heterogeneity} evaluated best practices for using the C+T method and found out that the optimal inclusion threshold is trait-dependent. 
[TOMOD] In this study, we particularly extend the work of \cite{abraham2013performance} by using a simulation framework similar but more extenstive than the one used in \cite{zhou2013polygenic}. 
In order to make our comparison as comprehensive as possible, we compare different disease architectures by varying the number, size and location of causal effects as well as the heritability. We also compare different models for generating liability scores, one with only linear effects, and one which combines linear, dominant and interaction-type effects.

[TOMOD] First, we quickly discard T-Trees because this method shows lower predictive performance as compared to other methods, while being computationally demanding.
Then, we show that the penalized logistic regression consistently achieves higher predictive performance while being faster than the C+T method, whereas predictive performance of the C+T method is very sensitive to the threshold of inclusion of SNPs, depending on the disease architecture, as shown in \cite{ware2017heterogeneity}.

@@

We use real genotypes of European individuals from a case-control celiac disease study \cite[]{dubois2010multiple}. The composition of this dataset is presented in table \ref{tab:celiac-data}.
Details of quality control and imputation for this dataset are available in \cite{prive2017efficient}.
For simulations presented later, we first restrict the dataset to controls in order to remove the genetic structure induced by the celiac disease status. 
Then, we decided to remove population structure because it can affect the predictive performance of methods \cite[]{martin2017human}.
In order to alleviate population structure, we keep people from the UK only and we further remove outliers based on the robust Mahalanobis distances computed using the first 10 principal components of remaining individuals. 
This 3-step filtering process results in a sample of 7102 individuals with minimal population structure (see supplementary notebook ``preprocessing'').

@@

In order to simulate phenotypes, we use the Liability Threshold Model (LTM) with a prevalence of 30\% \cite[]{falconer1965inheritance}. We vary different parameters: the number of chromosomes used (either all 22 chromosomes or chromosome 6 only), the size of the training set (from 1000 to 6000), the number of causal variants and their location (30, 300 or 3000 anywhere on the genome or 30 only in the HLA region), the heritability (50\% or 80\%), the distribution of effects associated with causal SNPs (Normal or Laplace). We also use two different models to compute liability scores: a ``simple'' model with linear effects only, and a ``fancy'' model which combines linear, dominant and interaction-type effects.
For the ``simple'' model, we compute $$y_i = \sum_{j\in S_\text{causal}} w_j \cdot \widetilde{G_{i,j}}$$ where $w_j$ are weights generated with a Gaussian or a Laplace distribution, $G_{i,j}$ is the allele count of individual $i$ for SNP $j$ and $\widetilde{G_{i,j}}$ corresponds to its standardized version (zero mean and unit variance for all SNPs). 
For the ``fancy'' model, we separate the causal SNPs in three equal sets $S_\text{causal}^{(1)}$, $S_\text{causal}^{(2)}$ 
and $S_\text{causal}^{(3)}$; $S_\text{causal}^{(3)}$ is further separated in two equal sets, $S_\text{causal}^{(3.1)}$ and $S_\text{causal}^{(3.2)}$. 
We then compute $$y_i = \underbrace{\sum_{j\in S_\text{causal}^{(1)}} w_j \cdot \widetilde{G_{i,j}}}_\text{linear} + \underbrace{\sum_{j\in S_\text{causal}^{(2)}} w_j \cdot \widetilde{D_{i,j}}}_\text{dominant} + \underbrace{\sum_{\substack{k=1 \\ j_1=e_k^{(3.1)} \\ j_2=e_k^{(3.2)}}}^{k=\left|S_\text{causal}^{(3.1)}\right|} w_{j_1} \cdot \widetilde{G_{i,j_1} G_{i,j_2}}}_\text{interaction}$$ 
where $D_{i,j} = \mathds{1}\left\{G_{i,j} \neq 0\right\}$ and $S_\text{causal}^{(q)}=\left\{e_k^{(q)},~k \in \left\{1, \ldots, \left|S_\text{causal}^{(q)}\right|\right\}\right\}$. 
Note that for the interaction part of the model, we scale interactions, not the raw allele counts, so that corresponding SNPs still have a marginal effect. 
For both models, the values of $y_i$ are then standardized so that they have a variance equal to the desired heritability $h^2$ and we further add some environmental noise $\epsilon_i$ to $y_i$, where $\epsilon \sim N(0, 1 - h^2)$.

We implement 3 different simulation scenarios, summarized in table \ref{tab:simus}. 
Scenario \textnumero1 uses the whole dataset (all 22 autosomal chromosomes) and a training set of size 6000. 
It compares all methods described in section \ref{sec:methods}; for each combination of the remaining parameters, results are based on 100 simulations (i.e.\ 5 times more than in \cite{zhou2013polygenic}). 
However, comparison results for T-Trees relies on 5 simulations only (and an heritability of 80\%) because this method can require several hours of computation for a single simulation. 
Scenario \textnumero2 consists of 100 simulations per combination of parameters on a dataset composed of chromosome 6 only. Reducing the number of SNPs aims at increasing the polygenicity (the proportion of causal SNPs) of the simulated models and at virtually increasing the sample size \cite[]{dudbridge2013power,marquez2017multiethnic,vilhjalmsson2015modeling}. 
For this scenario, we use only the additive (``simple'') model, but continue to compare all previous different values of the other parameters.
Finally, scenario \textnumero3 reuses the whole dataset but this time varying the size of the training set in order to assess how the sample size affects predictive performance of methods. A total of 100 simulations per combination of parameters are run using 300 causal SNPs anywhere on the genome.

@@

In this study, we use two different measures of predictive accuracy. 
First, we use the Area Under the Receiver Operating Characteristic (ROC) Curve (AUC) \cite[]{fawcett2006introduction,hanley1982meaning,lusted1971signal}. In the case of our study, the AUC is the probability that a PRS of a case is greater than the PRS of a control.
This measure indicates the extent to which we can distinguish between cases and controls using PRSs.
As a second measure, the partial AUC for specificities between 90\% and 100\% is also reported \cite[]{dodd2003partial,mcclish1989analyzing}. This measure is similar to the AUC, but focuses on high specificities, which is the most useful part of the ROC curve in clinical settings.

When reporting results of AUC, we use estimates of maximum achievable AUC values for the simulations: 84\% and 94\% for heritabilities of respectively 50\% and 80\% based on three different estimations (see Supplementary Materials).

%Note that we also report the timing of the main computations and the number of SNPs used in the predictions.

@@

When there are 30 causal SNPs and 80\% of heritability, penalized logistic regression provides AUC greater than 93\%, nearly reaching the maximum achievable AUC of 94\%, whereas AUC values obtained with C+T method range between 83\% and 90\% (Figure \ref{fig:main-AUC-logit}). 
Moreover, the penalized logistic regression consistently provides better predictive performance than the C+T method for all disease architectures tested, excepted in some simulations with 3000 causal SNPs, which are cases where all methods hardly exceed 60\% of AUC (Figures \ref{fig:main-AUC-logit} and \ref{fig:supp-AUC-logit}).

@@

As for the C+T method itself, we recall that we report the maximum AUC of 102 thresholds, a threshold that should normally be learned on the training set only. The choice of this threshold is very important as it can greatly impact the predictive performance of the method \cite[]{ware2017heterogeneity}. 
On the contrary, in the penalized logistic regression presented here, we remove the choice of the regularization parameter with a procedure that we call Cross-Model Selection and Averaging (CMSA) while keeping high predictive performance [TODO: COMPARAISON AVEC LE MAX DE BIGLASSO]. Usually, cross-validation is used to assess the best value for the hyper-parameter and then the model is train again with this particular hyper-parameter value \cite[]{Hastie2008,wei2013large}. Yet, as the value of $\lambda$ depends on the sample size and doing cross-validation and retraining is computationally demanding, we prefer to use CMSA when there is a model to choose from a regularization path, with the possibility of adding an early stopping criterion to fasten the whole procedure.

We also introduce a simple feature engineering technique which enables the logistic regression to detect and learn not only additive effects, but also dominant and recessive ones [DEJA ETE FAIT? CITATION?]. 
This technique improves the predictive performance of the method when there are some non-linear effects in the simulations, while providing the same predictive performance when there are only linear effects. Moreover, it also improves predictive performance for the celiac disease.

Finally, in this comparative study, we did not consider the problem of population structure \cite[]{marquez2017multiethnic,martin2017human,vilhjalmsson2015modeling} and also did not consider non-genetic data such as environmental and clinical data \cite[]{dey2013integration,van2012integration}. In another study, we will assess how can we use models and effects learned in one population to improve learning and prediction in another population. We will use a dataset as large as the UK biobank \cite[]{bycroft2017genome}. In another study, we wish also to assess how can we combine the information provided by genetic data with clinical and environmental data, possibly in a non-linear way.

@@

>>>>

\title{Efficient implementation of penalized models and their application to genetic risk prediction}
\author{Florian Priv\'e\,$^{\text{ 1,}*}$, Hugues Aschard\,$^{\text{2}}$ and Michael G.B. Blum\,$^{\text{1,}*}$}

@@

\abstract{
Polygenic Risk Scores (PRS) consist in combining the information across many single-nucleotide polymorphisms (SNPs) in a score reflecting the genetic risk of developing a disease. PRSs might have a major public health impact, possibly allowing for screening campaigns to identify high-genetic risk individuals for a given disease.  The ``Clumping+Thresholding'' (C+T) approach, which is is the most common method to derive PRS, uses only univariate genome-wide association studies (GWAS) summary statistics, which makes it suitable and fast. However, previous work showed that jointly estimating SNP effects for computing PRS have the potential to significantly improve the predictive performance of PRS as compared to C+T.

In this paper, we present an efficient implementation of such penalized joint estimation, allowing for practical application of penalized logistic regression on modern datasets including hundreds of thousands of individuals. Our implementation of penalized logistic regression also includes nearly-optimal choices for hyper-parameters. The choice of hyper-parameters for a method is very important as it can dramatically impact the predictive performance of the method (e.g. the p-value threshold in C+T).

We compare the performance of penalized logistic regression to the C+T method and to a derivation of random forests. The penalized logistic regression consistently achieves higher predictive performance than the two other methods while being faster. We also found that improvement in predictive performance is more pronounced when there are few effects located in nearby genomic regions with correlated SNPs; AUC values increase from 83\% with C+T to 92.5\% with penalized logistic regression. We confirmed these results in a data analysis of a case-control study for celiac disease where penalized logistic regression and the standard C+T method achieve AUC of 89\% and of 82\%.

In conclusion, our study demonstrates that penalized logistic regression are applicable to large-scale individual-level data and can achieve more discriminative polygenic risk scores. Our implementation is publicly available in the R package bigstatsr.\\

@@

Many methods have been developed to predict disease status, or more generally any phenotype, based on SNP information. A commonly used method, called ``P+T'' or ``C+T'' (which stands for ``Clumping and Thresholding'') -- or even just PRS -- is used to derive PRSs from results of Genome-Wide Association Studies (GWASs) \cite[]{chatterjee2013projecting,dudbridge2013power,evans2009harnessing,purcell2009common,wray2007prediction}. This technique uses GWAS summary statistics only, allowing for trivial and fast implementation. However, the approach also has limitations. In particular, previous studies showed that predictive performance of the C+T method is very sensitive to the threshold of inclusion of SNPs, depending on the disease architecture \cite{ware2017heterogeneity}.
Linear Mixed-Models (LMMs) are another widely-used method in fields such as plant and animal breeding or for predicting highly heritable quantitative human phenotypes such as height \cite[]{lello2017accurate,yang2010common}. Yet, models resulting from LMM, known e.g.\ as ``gBLUP'', are not optimal for predicting disease status based on genotypes \cite[]{abraham2013performance}. Moreover, these methods and their derivatives are often computationally demanding, both in terms of memory and time required, which makes them unlikely to be used for prediction on very large datasets \cite[]{golan2014effective,maier2015joint,speed2014multiblup,zhou2013polygenic}.
Finally, statistical learning methods have also been used to derive PRSs for complex human diseases by jointly estimating SNP effects. Such methods include joint logistic regression, Support Vector Machine (SVM) and random forests \cite[]{abraham2012sparsnp,abraham2014accurate,botta2014exploiting,okser2014regularized,wei2009disease}.

We recently developed two R packages, bigstatsr and bigsnpr, for efficiently analyzing large-scale genome-wide data \cite[]{prive2017efficient}. Package bigstatsr includes an efficient algorithm with a new implementation for computing sparse linear and logistic regressions on huge datasets as large as the UK Biobank \cite[]{bycroft2017genome}.
In this paper, we present a comprehensive comparative study of our implementation of penalized logistic regression against the C+T method and the T-Trees algorithm, a derivation of random forests that has shown high predictive performance \cite[]{botta2014exploiting}.
Note that in this comparison we do not include any LMM method for the reasons mentioned before and do not include any SVM method because it is expected to give similar results to logistic regression \cite[]{abraham2012sparsnp}. For the C+T model, we compare different thresholds of inclusion of SNPs. For the penalized logistic regression, we also introduce two improvements. First, we present a procedure that we call Cross-Model Selection and Averaging (CMSA) for automatically choosing the regularization parameter of the penalized logistic regression, which directly affects the number of SNPs included in the model. Secondly, we also show how to use feature engineering in order to capture not only linear effects, but also recessive and dominant effects.

For the simulations of this paper, we use real genotype data and simulate new phenotypes \cite{zhou2013polygenic}. 
In order to make our comparison as comprehensive as possible, we compare different disease architectures by varying the number, size and location of causal effects as well as the heritability. We also compare different models for simulating new phenotypes, one with only linear effects, and one which combines linear, dominant and interaction-type effects.
Overall, we found that the penalized logistic regression consistently achieves higher predictive performance while being faster than the C+T and T-Trees methods. This demonstrates the feasibility and relevance of this approach in PRS modeling in future studies.

@@

We use real genotypes of European individuals from a case-control celiac disease study \cite[]{dubois2010multiple}. The composition of this dataset is presented in table \ref{tab:celiac-data}. Details of quality control and imputation for this dataset are available in \cite{prive2017efficient}. For simulations presented later, we first restrict the dataset to controls in order to remove the genetic structure induced by the celiac disease status. Then, we decided to remove population structure because it can affect the predictive performance of methods \cite[]{martin2017human}.In order to alleviate population structure, we keep people from the UK only and we further remove outliers based on the robust Mahalanobis distances computed using the first 10 principal components of remaining individuals. This 3-step filtering process results in a sample of 7102 individuals with minimal population structure (see supplementary notebook ``preprocessing''). We also use this dataset for a real data application, in this case keeping all individuals.

@@

We simulate binary phenotypes using a Liability Threshold Model (LTM) with a prevalence of 30\% \cite[]{falconer1965inheritance}. We vary parameters of the simulations in order to match a range of genetic architecture, from low to high polygenicity. For practical reason, this was achieved by varying the number of chromosomes used (either all 22 autosomal chromosomes or chromosome 6 only), the number of causal variants and their location (30, 300 or 3000 anywhere on the genome or 30 only in the HLA region), and the heritability (50\% or 80\%). We also considered deviation form the standard normal additive model, drawing effects of causal SNPs from either Normal or Laplace distribution, and computing liability scores either from a ``simple'' model with linear effects only or a ``fancy'' model which combines linear, dominant and interaction-type effects. For the ``simple'' model, we compute $$y_i = \sum_{j\in S_\text{causal}} w_j \cdot \widetilde{G_{i,j}} ~~+~~ \epsilon_i$$ where $w_j$ are weights generated with a Gaussian or a Laplace distribution, $G_{i,j}$ is the allele count of individual $i$ for SNP $j$, $\widetilde{G_{i,j}}$ corresponds to its standardized version (zero mean and unit variance for all SNPs) and $\epsilon \sim N(0, 1 - h^2)$.
For the ``fancy'' model, we separate the causal SNPs in three equal sets $S_\text{causal}^{(1)}$, $S_\text{causal}^{(2)}$
and $S_\text{causal}^{(3)}$; $S_\text{causal}^{(3)}$ is further separated in two equal sets, $S_\text{causal}^{(3.1)}$ and $S_\text{causal}^{(3.2)}$.
We then compute $$y_i = \underbrace{\sum_{j\in S_\text{causal}^{(1)}} w_j \cdot \widetilde{G_{i,j}}}_\text{linear} + \underbrace{\sum_{j\in S_\text{causal}^{(2)}} w_j \cdot \widetilde{D_{i,j}}}_\text{dominant} + \underbrace{\sum_{\substack{k=1 \\ j_1=e_k^{(3.1)} \\ j_2=e_k^{(3.2)}}}^{k=\left|S_\text{causal}^{(3.1)}\right|} w_{j_1} \cdot \widetilde{G_{i,j_1} G_{i,j_2}}}_\text{interaction} ~~+~~ \epsilon_i$$
where $D_{i,j} = \mathds{1}\left\{G_{i,j} \neq 0\right\}$ and $S_\text{causal}^{(q)}=\left\{e_k^{(q)},~k \in \left\{1, \ldots, \left|S_\text{causal}^{(q)}\right|\right\}\right\}$.
Note that, because we used the raw genotype counts when coding the interaction terms, the corresponding SNPs also display a marginal effect.

We implement 3 different simulation scenarios, summarized in table \ref{tab:simus}.
Scenario \textnumero1 uses the whole dataset (all 22 autosomal chromosomes) and a training set of size 6000. It compares all methods described in section \ref{sec:methods}; for each combination of the remaining parameters, results are based on 100 simulations except  for T-Trees, which relies on 5 simulations only (and an heritability of 80\%) because of a much higher computational burden (several hours of computation for a single simulation) as compared to other approaches. Scenario \textnumero2 consists of 100 simulations per combination of parameters on a dataset composed of chromosome 6 only. Reducing the number of SNPs aims at increasing the polygenicity (the proportion of causal SNPs) of the simulated models and at virtually increasing the sample size \cite[]{dudbridge2013power,marquez2017multiethnic,vilhjalmsson2015modeling}.
For this scenario, we use only the additive (``simple'') model, but continue to compare all previous different values of the other parameters. Finally, scenario \textnumero3 reuses the whole dataset but this time varying the size of the training set in order to assess how the sample size affects predictive performance of methods. A total of 100 simulations per combination of parameters are run using 300 causal SNPs anywhere on the genome.

@@

In this study, we use two different measures of predictive accuracy. First, we use the Area Under the Receiver Operating Characteristic (ROC) Curve (AUC) \cite[]{fawcett2006introduction,lusted1971signal}. In the case of our study, the AUC is the probability that a PRS of a case is greater than the PRS of a control.
This measure indicates the extent to which we can distinguish between cases and controls using PRSs. As a second measure, the partial AUC for specificities between 90\% and 100\% is also reported \cite[]{dodd2003partial,mcclish1989analyzing}. This measure is similar to the AUC, but focuses on high specificities, which is the most useful part of the ROC curve in clinical settings. When reporting results of AUC, we use estimates of maximum achievable AUC values for the simulations: 84\% and 94\% for heritabilities of respectively 50\% and 80\% based on three different estimations (see Supplementary Materials). %Note that we also report the timing of the main computations and the number of SNPs used in the predictions.

@@

When there are 30 causal SNPs and 80\% of heritability, penalized logistic regression provides AUC greater than 93\%, nearly reaching the maximum achievable AUC of 94\%, whereas AUC values obtained with C+T method range between 83\% and 90\% (Figure \ref{fig:main-AUC-logit}). Moreover, the penalized logistic regression consistently provides better predictive performance than the C+T method across all scenarios we considered, except in the case of high polygenicity and small sample size, in which case all methods hardly exceed 60\% of AUC (Figures \ref{fig:main-AUC-logit} and \ref{fig:supp-AUC-logit}).

@@

As for the C+T method itself, we recall that we report the maximum AUC of 102 thresholds, a threshold that should normally be learned on the training set only. The choice of this threshold is very important as it can greatly impact the predictive performance of the method \cite[]{ware2017heterogeneity}. On the contrary, in the penalized logistic regression presented here, we remove the choice of the regularization parameter with a procedure that we call Cross-Model Selection and Averaging (CMSA) while keeping high predictive performance [TODO: COMPARAISON AVEC LE MAX DE BIGLASSO]. Usually, cross-validation is used to assess the best value for the hyper-parameter and then the model is train again with this particular hyper-parameter value \cite[]{Hastie2008,wei2013large}. Yet, as the value of $\lambda$ depends on the sample size and doing cross-validation and retraining is computationally demanding, we prefer to use CMSA when there is a model to choose from a regularization path, with the possibility of adding an early stopping criterion to fasten the whole procedure. We also introduce a simple feature engineering technique which enables the logistic regression to detect and learn not only additive effects, but also dominant and recessive ones. This technique improves the predictive performance of the method when there are some non-linear effects in the simulations, while providing the same predictive performance when there are only linear effects. Moreover, it also improves predictive performance for the celiac disease.Finally, in this comparative study, we did not consider the problem of population structure \cite[]{marquez2017multiethnic,martin2017human,vilhjalmsson2015modeling} and also did not consider non-genetic data such as environmental and clinical data \cite[]{dey2013integration,van2012integration}. In another study, we will assess how can we use models and effects learned in one population to improve learning and prediction in another population. We will use a dataset as large as the UK biobank \cite[]{bycroft2017genome}. In another study, we wish also to assess how can we combine the information provided by genetic data with clinical and environmental data, possibly in a non-linear way.

@@

\newpage
\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{supp-biglasso}}
\caption{Comparison of ``logit-triple'' and the best prediction for ``biglasso'' (another implementation of penalized logistic regression without procedure for choosing the $\lambda$ hyper-parameter) in scenario \textnumero1. Vertical panels are presenting results for effects following a Gaussian or Laplace distribution. Horizontal panels are presenting results for the ``simple'' model for simulating phenotypes. Simulations use an heritability of 80\%. \textbf{A:} Mean of AUC over 100 simulations. Error bars are representing $\pm 2 \text{SD}$ of $10^5$ non-parametric bootstrap of the mean of AUC. The blue dotted line represents the maximum achievable AUC. \textbf{B:} Boxplots of numbers of predictors used by the methods for 100 simulations. \textbf{C:} Boxplots of execution times for 100 simulations.}
\label{fig:supp-biglasso}
\end{figure}

\newpage
\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{supp-alpha}}
\caption{Comparison of different values of the $\alpha$ hyper-parameter of the ``logit-simple'' method in scenario \textnumero1. ``logit-simple'', ``logit-simple-enet-ridge'' and ``logit-simple-lasso'' corresponds to values of 0.5, 0.001 and 1, respectively. Vertical panels are presenting results for effects following a Gaussian or Laplace distribution. Horizontal panels are presenting results for the ``simple'' model for simulating phenotypes. Simulations use an heritability of 80\%. \textbf{A:} Mean of AUC over 100 simulations. Error bars are representing $\pm 2 \text{SD}$ of $10^5$ non-parametric bootstrap of the mean of AUC. The blue dotted line represents the maximum achievable AUC. \textbf{B:} Boxplots of numbers of predictors used by the methods for 100 simulations. \textbf{C:} Boxplots of execution times for 100 simulations.}
\label{fig:supp-alpha}
\end{figure}
